---
title: "Hmk_05"
author: Beatrice
format: html
editor: visual
---

Please read the entire [R for Data Science](https://r4ds.had.co.nz/transform.html) before you do this homework.

This homework relies on the `nycflights13` package, which contains several data frames, including `airlines`, `airports`, `flights`, `planes`, and `weather`. Loading `nycflights13` puts all of these data frames on the search path, so that they are available automatically, just like `mpg` or `mtcars`.

## Installing data packages

Remember that any package needs to be installed only once (per version of R), but needs to be loaded every time we start a new R session.

To install `nyclflights13`, we use `install.packages("nycflights13")` (with quotation marks). To load it, we use `library(nycflights13)`.

-   **I will first install the package and then load it. Then I will glimpse into it**

```{r, message=FALSE}
library(nycflights13)
```

```{r, message=FALSE}
library(tidyverse)
glimpse(flights)
```

# Question 1: filtering

Make a plot of air time as a function of distance for all flights that meet the following criteria:

-   originate from LaGuardia airport ("LGA")

-   departed on the 16th of the month

-   have a flight distance of less than 2000

-   **I am first going to assign all these filters to a specific object. They will all be combined into a single object.**

```{r, message=FALSE}
total <- filter(flights, origin == "LGA", day == 16, distance < 2000)
```

-   **Now I will make the plot as with the criteria stated above**

```{r, message=FALSE}
ggplot(total, aes(x=distance, y=air_time)) +
  geom_point()
```

# Question 2: dealing with NAs

Make a data frame of all of the rows of `flights` that have values for *both* `arr_time` and `dep_time` - that is, neither of those values are `NA`.

```{r, message=FALSE}
ques_2 <- filter(flights, arr_time > 0, dep_time > 0)
```

## filtering NAs

`ggplot()` will automatically remove NA values from the plot, as you may have seen in question 1, but it emits a warning message about that. Of course you could silence the warning message using [chunk options](https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html), but how could you prevent them from appearing in the first place?

* **If you use drop_na(), it should get rid of all the NAs in your data frame.**

# Question 3: adding columns

Create a data frame of average flight speeds, based on `air_time` and `distance`. Make either a histogram or a density plot of the data. If you like, you may break the data out (e.g. by airline, or some other variable) in a way that you think makes sense.

```{r, message=FALSE}
ques_3 <- mutate(flights, speed = distance/air_time)

ggplot(ques_3, aes(x=speed)) +
  geom_histogram(color="hotpink")
```